---
title: "Business Intelligence Lab Submission Markdown"
author: "naive"
date: "4/10/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  135575 - B - Dennis Nzioki.                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  134645 - B - Vivean Lydiah                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  134765 - B - Nicholas Munene                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  131653- B - Terry Joan                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 5.  124428 - B - Eston Gichuhi                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | *\<specify the name of the team you created on GitHub classroom\>*                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages
We start by installing all the required packages
```{r setup}
if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## naniar ----
if (require("naniar")) {
  require("naniar")
} else {
  install.packages("naniar", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## ggplot2 ----
if (require("ggplot2")) {
  require("ggplot2")
} else {
  install.packages("ggplot2", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## corrplot ----
if (require("corrplot")) {
  require("corrplot")
} else {
  install.packages("corrplot", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## ggcorrplot ----
if (require("ggcorrplot")) {
  require("ggcorrplot")
} else {
  install.packages("ggcorrplot", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
## dplyr ----
if (require("dplyr")) {
  require("dplyr")
} else {
  install.packages("dplyr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
```
# Step2 Load the dataset
```{r}
# load the dataset PimaIndians Diabetes2 from mlbench package
data("PimaIndiansDiabetes2")

PimaIndiansDiabetes2$diabetes <- factor(PimaIndiansDiabetes2$diabetes)

str(PimaIndiansDiabetes2)
dim(PimaIndiansDiabetes2)
head(PimaIndiansDiabetes2)
summary(PimaIndiansDiabetes2)
```
# Step3: check for any missing data and address it
```{r}
# Are there missing values in the dataset?
any_na(PimaIndiansDiabetes2)

# How many?
n_miss(PimaIndiansDiabetes2)

# What is the proportion of missing data in the entire dataset?
prop_miss(PimaIndiansDiabetes2)

# What is the number and percentage of missing values grouped by
# each variable?
miss_var_summary(PimaIndiansDiabetes2)

# Which variables contain the most missing values?
gg_miss_var(PimaIndiansDiabetes2)

# Which combinations of variables are missing together?
gg_miss_upset(PimaIndiansDiabetes2)

# Where are missing values located (the shaded regions in the plot)?
vis_miss(PimaIndiansDiabetes2) +
  theme(axis.text.x = element_text(angle = 80))
PimaIndiansDiabetes2_removed_obs <-
  PimaIndiansDiabetes2 %>%
  dplyr::filter(complete.cases(.))
## OPTION 1: Remove the observations with missing values ----
# We can decide to remove all the observations that have missing values
# as follows:
PimaIndiansDiabetes2_removed_obs <- PimaIndiansDiabetes2 %>% filter(complete.cases(.))

# The initial dataset had 768 observations and 9 variables
dim(PimaIndiansDiabetes2)

# The filtered dataset has 392 observations and 9 variables
dim(PimaIndiansDiabetes2)

# Are there missing values in the dataset?
any_na(PimaIndiansDiabetes2_removed_obs)
```
# STEP 4. Perform EDA and Feature Selection ----
## Compute the correlations between variables ----
We identify the correlated variables because it is these correlated variables
that can then be used to identify the clusters.
```{r}
# Create a correlation matrix
# Option 1: Basic Table
cor(PimaIndiansDiabetes2_removed_obs[, c(1,2,3,4,5,6,7,8)]) %>%
  View()

# Option 2: Basic Plot
cor(PimaIndiansDiabetes2_removed_obs[, c(1,2,3,4,5,6,7,8)]) %>%
  corrplot(method = "square")

# Option 3: Fancy Plot using ggplot2
corr_matrix <- cor(PimaIndiansDiabetes2_removed_obs[, c(1,2,3,4,5,6,7,8)])

p <- ggplot2::ggplot(data = reshape2::melt(corr_matrix),
                     ggplot2::aes(Var1, Var2, fill = value)) +
  ggplot2::geom_tile() +
  ggplot2::geom_text(ggplot2::aes(label = label_wrap(label, width = 10)),
                     size = 4) +
  ggplot2::theme_minimal() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

ggcorrplot(corr_matrix, hc.order = TRUE, type = "lower", lab = TRUE)

# The correlation plot shows a -0.02 correlation between the pedigree  and the
# pressure This is worth investigating further if the intention
# of the model is to diagnose diabetes based on pedigree.

# Room_type, neighbourhood, date and other non-numeric variables and
# categorical variables are not included in the correlation, but they can be
# used as an additional dimension when plotting the scatter plot during EDA.

## Plot the scatter plots ----
# A scatter plot to show pedigree levels against pressure
# per diabetes level
ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(pedigree, pressure,
           color = diabetes,
           shape = diabetes)) +
  geom_point(alpha = 0.5) +
  xlab("pedigree") +
  ylab("pressure")

# A scatter plot to show pedigree levels against pressure
# per age
ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(pedigree, pressure,
           color = age)) +
  geom_point(alpha = 0.5) +
  xlab("pedigree") +
  ylab("pressure")

# A scatter plot to show insulin levels against age
# per diabetes level
ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(insulin, age,
           color = diabetes,
           shape = diabetes)) +
  geom_point(alpha = 0.5) +
  xlab("Insulin level") +
  ylab("age")

# A scatter plot to show glucose level against age 
#per diabetes levels
ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(glucose, age,
           color = diabetes,
           shape = diabetes)) +
  geom_point(alpha = 0.5) +
  xlab("glucose level") +
  ylab("age")

# A scatter plot to show mass against insulin levels
#  per diabetes type
ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(mass, insulin,
           color = diabetes,
           shape = diabetes)) +
  geom_point(alpha = 0.5) +
  xlab("mass") +
  ylab("Insulin level")

## Transform the data ----
# The K Means Clustering algorithm performs better when data transformation has
# been applied. This helps to standardize the data making it easier to compare
# multiple variables.

summary(PimaIndiansDiabetes2_removed_obs)
model_of_the_transform <- preProcess(PimaIndiansDiabetes2_removed_obs,
                                     method = c("scale", "center"))
print(model_of_the_transform)
PimaIndiansDiabetes2_removed_obs_std <- predict(model_of_the_transform, # nolint
                                                PimaIndiansDiabetes2_removed_obs)
summary(PimaIndiansDiabetes2_removed_obs_std)
sapply(PimaIndiansDiabetes2_removed_obs_std[, c(1, 2, 3, 4, 5, 6, 7, 8)], sd)

## Select the features to use to create the clusters ----
# OPTION 1: Use all the numeric variables to create the clusters
PimaIndiansDiabetes2_vars <-
  PimaIndiansDiabetes2_removed_obs_std[, c(1, 2, 3, 4, 5, 6, 7, 8)]
```
#Step5 Create the clusters using the K-Means Clustering Algorithm
```{r}
# We start with a random guess of the number of clusters we need
set.seed(7)
kmeans_cluster <- kmeans(PimaIndiansDiabetes2_vars, centers = 3, nstart = 20)

# We then decide the maximum number of clusters to investigate
n_clusters <- 8

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(7)

# Investigate 1 to n possible clusters (where n is the maximum number of 
# clusters that we want to investigate)
for (i in 1:n_clusters) {
  # Use the K Means cluster algorithm to create each cluster
  kmeans_cluster <- kmeans(PimaIndiansDiabetes2_vars, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- kmeans_cluster$tot.withinss
}

## Plot a scree plot ----
# The scree plot should help you to note when additional clusters do not make
# any significant difference (the plateau).
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)

scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
  geom_point(size = 4) +
  geom_line() +
  scale_x_continuous(breaks = c(2, 4, 6, 8)) +
  xlab("Number of Clusters")

scree_plot

# We can add guides to make it easier to identify the plateau (or "elbow").
scree_plot +
  geom_hline(
    yintercept = wss,
    linetype = "dashed",
    col = c(rep("#000000", 5), "#FF0000", rep("#000000", 2))
  )

# The plateau is reached at 6 clusters.
# We therefore create the final cluster with 6 clusters
# (not the initial 3 used at the beginning of this STEP.)
k <- 6
set.seed(7)
# Build model with k clusters: kmeans_cluster
kmeans_cluster <- kmeans(PimaIndiansDiabetes2_vars, centers = k, nstart = 20)

# STEP 6. Add the cluster number as a label for each observation ----
PimaIndiansDiabetes2_removed_obs$cluster_id <- factor(kmeans_cluster$cluster)

## View the results by plotting scatter plots with the labelled cluster ----
ggplot(PimaIndiansDiabetes2_removed_obs, aes(pedigree, pressure,
                                         color = cluster_id)) +
  geom_point(alpha = 0.5) +
  xlab("pedigree") +
  ylab("pressure")

ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(age, insulin, color = cluster_id)) +
  geom_point(alpha = 0.5) +
  xlab("age") +
  ylab("Insulin level")

ggplot(PimaIndiansDiabetes2_removed_obs,
       aes(insulin, pregnant,
           color = cluster_id)) +
  geom_point(alpha = 0.5) +
  xlab("Insulin level") +
  ylab("Pregnancy level")
```
